Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-12 MasterThesis.tex
Updates from Overleaf

%Title parameters
\title{Capturing underlying musical structure: an audio-based approach using deep learning}
\author{\href{oriol.colome.font@epidemicsound.com}{Oriol Colomé Font}}
\submitdate{July 2023}
\supervisor{Carlos Lordelo and Carl Thomé}
\cosupervisor{Frederic Font}

@@

%\maketitle
\restoregeometry

@@

%Introduction of the project
\input{introduction/introduction}
%Music analysis explanation
\input{music analysis/MSA}
%Methods
\input{methods/methods}
%Results
\input{results/results}
%Discussion
\input{discussion/discussion}

>>>>

%Title parameters
\title{%
  Uncovering underlying musical content in the temporal domain \\
  \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks}

\author{\href{oriol.colome.font@epidemicsound.com}{Oriol Colomé Font}}
\submitdate{July 2023}
\supervisor{Carlos Lordelo and Carl Thomé}
\cosupervisor{Frederic Font}

@@

\restoregeometry

@@

%Introduction of the project
\input{introduction/introduction}
%Task definition
%%\input{task definition}
%Methods
\input{methods/methods}
%Results
\input{results/results}
%Discussion
\input{discussion/discussion}

### 2023-03-12 introduction/Subsections/Objectives.tex
Updates from Overleaf

As a professional music transcriber myself, I aim to understand how deep neural networks (DNNs) filter musical features, listen to the NN layers from a musician's perspective and analyze the resulting data. I aim to approach the technology from a technical and musical standpoint to understand how DNNs process and extract musical information. By doing so, I aim to enhance the accuracy and efficiency of the process and contribute to developing AI technology for music analysis. By combining my expertise in music with my -not fully developed- knowledge of DNNs, I hope to bring a unique perspective to the field and contribute to the advancement of music analysis.

Music is a complex and multi-faceted phenomenon, and implementing a model that mimics human aural music skills is challenging. However, using the latest deep learning techniques, we aim to build a model that achieves good results for a specific task in the domain of music.

>>>>

This work aims to generalize the fundamental tonal-working structure of western-tradition music by learning and embedding space used to facilitate its application in downstream MIR tasks. Understanding how music is organized and composed is critical, and music theory plays a vital role.

In recent years, using neural networks has enabled low-dimensional musical embeddings that capture crucial musical features and allow the computation of content-based similarity between musical passages. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, we can measure the similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure the output aligns with human expectations.

This study aims to comprehensively analyze musical works from diverse styles and genres by incorporating computational techniques from Music Information Retrieval (MIR) and a musicological perspective to contextualize the findings. Aural skills are required to comprehend musical works through listening, as they involve perceptual and cognitive abilities. The underlying musical structure refers to the systematic arrangement of elements that produce a coherent and expressive musical work. This study aims to offer fresh insights into these components of musical composition and advance our understanding of their relationships.

Music is a complex and multi-faceted phenomenon, and implementing a model mimicking human aural music skills is challenging. However, the latest deep learning techniques provide a promising avenue to build a model that achieves good results for a specific task in music. Therefore, this study will utilize deep learning techniques to investigate the correlation between aural skills and the underlying tonal-working structure of music.

Furthermore, this study will argue that MIR research should shift from a purely technical-centered approach to a more balanced and comprehensive one incorporating other relevant domains such as musicology and music theory. By doing so, we can better understand the interplay between technical and musical aspects of music and improve our understanding of music signals and the music itself.
